detect patches

### 2024-01-24 scripts/classify_patches.Rmd
class metrics to test

```{r}
p <- '../results/suitability_maps'
r <- rast(file.path(p, 'Achnatherum_aridum1k-2024-01-10_09:16:55.tif'))
msk <- ifel(r < 0.80, NA, r)
m <- mask(r, msk)
m1 <- aggregate(m, 3, fun = 'mean')

@@

@@

# euclidean nearest neighbor distance (enn)
## (modification) log(x), sqrt(x)
### Expectation: the number of colonized predicted patches decrease with increasing distance.
#### why: long-distance dispersal is limited from colonized to suitable patches. 
##### test: logistic regression, independent: distance, dependent: colonization (0|1)

sqrt(c( 1000, 150))
```

```{r Patch metrics to predict abundance}
library(landscapemetrics)

# Patch level metrics. 
print(
  list_lsm(level = "patch"), 
  n = 12)

>>>>

```{r}
p <- '../results/suitability_maps'
r <- rast(file.path(p, 'Achnatherum_aridum1k-2024-01-10_09:16:55.tif'))
msk <- ifel(r < 0.80, NA, r)
m <- mask(r, msk)
m1 <- aggregate(m, 2, fun = 'mean')

@@

aggregation by a factor of 2 allows for quick patch delineation and detection.

@@

# euclidean nearest neighbor distance (enn) - (Patch)
## (modification) log(enn), sqrt(enn)
### Expectation: the number of colonized predicted patches decrease with increasing distance.
#### why: long-distance dispersal is limited from colonized to suitable patches. 
##### test: glm(colonization(0|1) ~ enn,  family = binomial)

# nearest neighbor distance (enn_mn) - (Class) # note other method of aggregation can be used too!
## (modification) log(enn_mn), sqrt(enn_mn)
### Expectation: the number of colonized predicted patches increases with increasing distance.
#### why: this species is adept at long distance dispersal
##### test: glm(colonization(0|1) ~ enn_mn,  family = binomial)
```

```{r Patch metrics to predict abundance}
# library(landscapemetrics)

# Patch level metrics. 
print(
  landscapemetrics::list_lsm(level = "class"))
